### 2022-04-24 greta/hopf_mcmc.Rmd
stoch

```{r message=FALSE}
library(tidyverse, quietly = TRUE)
library(greta, quietly = TRUE)
library(bayesplot, quietly = TRUE)
source("R/utils.R")
source("R/hopf.R")
```

@@

```{r}
combined <- 
  compare_forecast(draws, train, test, simulate, vars = c("H", "P")) |> 
  pivot_longer(c(H, P), values_to="value", names_to="variable")

@@

```{r}
scores <- bind_rows(
  rep_scores(combined, "P"),
  rep_scores(combined, "H") |> 
  mutate(scenario="hopf_increasing", model="MCMC") 
)
```

@@

```{r}
combined <- 
  compare_forecast(draws2, train, test, simulate, vars = c("H", "P")) |> 
  pivot_longer(c(H, P), values_to="value", names_to="variable")
write_csv(combined, "data/hopf_decreasing.csv.gz")

@@

```{r}

scores <- bind_rows(
  rep_scores(combined, "P"),
  rep_scores(combined, "H") |> 
  mutate(scenario="hopf_decreasing", model="MCMC") 
) |> bind_rows(scores)

>>>>

```{r setup, message=FALSE}
library(tidyverse, quietly = TRUE)
library(greta, quietly = TRUE)
library(bayesplot, quietly = TRUE)
source("R/utils.R")
source("R/hopf.R")
```

@@

```{r}
combined <-  compare_forecast(draws, train, test, simulate, vars = c("H", "P"))

@@

```{r}
scores <- 
  bind_rows(
    rep_scores(combined, "P"),
    rep_scores(combined, "H") 
  ) |> 
  mutate(scenario="hopf_increasing", model="MCMC")
```

@@

```{r}
combined <- compare_forecast(draws2, train, test, simulate, vars = c("H", "P"))

write_csv(combined, "data/hopf_decreasing.csv.gz")

@@

```{r}
scores <- 
  bind_rows(
    rep_scores(combined, "P"),
    rep_scores(combined, "H") 
  ) |> 
  mutate(scenario="hopf_decreasing", 
         model="MCMC", 
         reps = train_reps) |>
  bind_rows(scores)

write_csv(scores, "data/scores_hopf.csv.gz")

### 2022-04-24 greta/stochastic_mcmc.Rmd
stoch

```{r}
library(tidyverse)
library(greta)
library(bayesplot)

set.seed(4242)
train_reps <- 1
train_t_max <- 250
test_t_max <- 250
test_reps <- 100

@@

```{r}

```{r}

train <- purrr::map_dfr(1:train_reps, \(i) simulate(t_max=train_t_max), .id = "i")
test <- purrr::map_dfr(1:test_reps, \(i) simulate(t_max=test_t_max), .id = "i")

## show training data
train |> ggplot(aes(t, N, group=i)) + geom_line(alpha=0.2)
```

```{r}

m <- greta_model_stoch(train)
```

```{r}
mmcmc <- memoise::memoise(mcmc, cache = memoise::cache_filesystem("mcmc_cache"))

bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 60000, warmup = 50000, chains = 4, verbose = FALSE)
})
```

@@

```{r}
## draw test_reps number of samples

posterior_samples <- 
  bind_rows(map(draws, as_tibble)) %>% 
  sample_n(test_reps)
posterior_sims <- posterior_samples %>%
  purrr::transpose() %>%
  map_dfr(function(q) simulate(t_max = test_t_max, p = q) ,.id = "i")

bind_rows(
  mutate(test, model="true"), 
  mutate(posterior_sims, model="predicted")
)|> 
  ggplot(aes(t, N, col=model, group=interaction(model,i))) +
           geom_line(alpha=0.1)

```{r}
true <- tibble(a = 0.75, b = 0.25, sigma = 1e-2) |> gather(variable, value)
bind_rows(map(draws, as_tibble)) |>
  gather(variable, value) |> ggplot() + 
  geom_histogram(aes(value), bins = 30)  +
  geom_vline(data = true, aes(xintercept = value), col = "red", lwd = 1) + 
  facet_wrap(~variable, scales = "free")
```

# Scoring

First, a single 'observed' sample from the true:

```{r}
library(scoringRules)

scores <- function(observed, dat) {
  logsscore <- scoringRules::logs_sample(observed, dat)
  crpsscore <- scoringRules::crps_sample(observed, dat)
  data.frame(logs = mean(logsscore[-1]), crps =  mean(crpsscore[-1]))

}
# ensemble predictions
dat <- 
  posterior_sims |> 
  pivot_wider(id_cols = "t", names_from="i", values_from = "N") |> 
  select(-t) |> as.matrix()

```
```{r}
# score over all replicates:
bench::bench_time({
rep_scores <- 
  test |> 
  group_by(i) |> 
  group_map(~ scores(.x$N, dat)) |> 
  bind_rows()
})
```

```{r}
rep_scores |> summarise(across(.fns= base::mean))

library(patchwork)
ggplot(rep_scores) + geom_histogram(aes(crps), bins=20) + 
(ggplot(rep_scores) + geom_histogram(aes(logs), bins=20) )

>>>>

```{r setup, message=FALSE}
library(tidyverse, quietly = TRUE)
library(greta, quietly = TRUE)
library(bayesplot, quietly = TRUE)
source("R/utils.R")
source("R/hopf.R")
```

```{r}
set.seed(4242)
train_reps <- 1
train_t_max <- 250
test_t_max <- 250
test_reps <- 100
simulate <- simulate_stoch
```

```{r}
train <- purrr::map_dfr(1:train_reps, \(i) simulate(t_max=train_t_max), .id = "i")
test <- purrr::map_dfr(1:test_reps, \(i) simulate(t_max=test_t_max), .id = "i")
```

```{r}
m <- greta_model_stoch(train)
bench::bench_time({                 
  draws <- mmcmc(m, n_samples = 60000, warmup = 50000,
                 chains = 4, verbose = FALSE)
})
```

```{r}
## draw test_reps number of samples
combined <- compare_forecast(draws, train, test, simulate, vars = "N") 
write_csv(combined, "data/stochastic.csv.gz")
```

```{r}
scores <-
  rep_scores(combined, "N") |> 
  mutate(scenario="stochastic", 
         model="MCMC", 
         reps = train_reps)

write_csv(scores, "data/scores_stochastic.csv.gz")

@@

@@

```{r}
combined |> 
  group_by(t,type,variable) |> 
  summarise(mean = mean(value), sd = sd(value), .groups = "drop") |> 
 ggplot(aes(t, col=type)) + 
  geom_ribbon(aes(ymin = mean-2*sd, ymax = mean+2*sd, fill=type), alpha=0.5) +
  geom_line(aes(y=mean)) +
  geom_vline(aes(xintercept = train_t_max)) + facet_wrap(~variable, ncol=1)
```

```{r}
single <- 
  combined |> 
  filter(i %in% 1:train_reps) |> 
  ggplot(aes(t, value, col=type, group=interaction(i,type))) + 
    geom_line() +
    geom_vline(aes(xintercept = train_t_max)) + facet_wrap(~variable, ncol=1)
```
